---
title: "Growth Yield Figures"
author: "Juan P Barraza"
date: "1/28/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(plotrix)#this allows the std.error function to work
library(cowplot)

#this is co-culture data from PA14 using exponential phase cultures
coCultureGCFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/PA14 exp growth curves.csv"
coCultureGCData <-read.csv(coCultureGCFile)
coCultureGCData%>% group_by(condition, bug, time, strain) %>% summarise(cellCount = mean(countsmL), cellSE = std.error(countsmL),  cellSD = sd(countsmL)) %>% as.data.frame() -> coCultureGCDataStats
levels(coCultureGCDataStats$condition)<-c("Stirred", "Static")

##this data is from non-exponential phase cultures, but the yield should be fine for the mono-cultures
#Data from PAO1
nonExpPAO1File<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/PAO1 growth curves.csv"
nonExpPAO1Data <-read.csv(nonExpPAO1File)
nonExpPAO1Data<- na.omit(nonExpPAO1Data)
nonExpPAO1Data$countsmL<-as.numeric(levels(nonExpPAO1Data$countsmL))[nonExpPAO1Data$countsmL]
nonExpPAO1Data %>% filter(condition=="Mono")->nonExpPAO1DataMono
#Data from PA14
nonExpPA14File<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/PA14 growth curves.csv"
nonExpPA14Data <-read.csv(nonExpPA14File)
nonExpPA14Data<- na.omit(nonExpPA14Data)
nonExpPA14Data %>% filter(condition=="Mono")->nonExpPA14DataMono

nonExpAllData<-rbind(nonExpPAO1DataMono,nonExpPA14DataMono)
nonExpAllData %>% filter(bug=="Sa")->nonExpAllDataSa#All non-exponential data from Sa
nonExpAllData %>% filter(!bug=="Sa")->nonExpAllDataPa#All non-exponential data from Pa

```
These are the growth curves for non exponential phase Pa and Sa mono culture growth curves. The growth yield should not be affected as cultures are still approaching carrying capacity since they are not being outcompeted by Pa. There are two data points significantly lower than the other 3.
```{r growth curves nonExp, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}
##Graph growth yields for non-exponential data
#First for Sa:
ggplot(nonExpAllDataSa, aes(x=time, y = countsmL, color = bug))+
  geom_jitter(height = 0, width = .2)+
  scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  #scale_x_continuous(breaks = c(0,4,8,12,16))+
  ggtitle("SCFM2 NonExp Sa mono growth")

#Then for Pa:
ggplot(nonExpAllDataPa, aes(x=time, y = countsmL, color = bug))+
  geom_jitter(height = 0, width = .2)+
  scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  #scale_x_continuous(breaks = c(0,4,8,12,16))+
  ggtitle("SCFM2 NonExp Pa mono growth")

```


This is the data yield starting from exponential phase and going up to 16 hours:

```{r exponential phase yield, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}
##Lines below are to go from 12hr growth curve to growth yield
#First for Sa:
nonExpAllDataSa %>% filter(time=="12")->nonExpAllDataSaYield
nonExpAllDataSaYield %>% select(countsmL,rep,bug) %>% mutate(condition = "static")->nonExpAllDataSaYieldSimple
nonExpAllDataSaYieldSimple<-nonExpAllDataSaYieldSimple[c(1,2,4,3)]
colnames(nonExpAllDataSaYieldSimple) <- c("bugs","rep", "condition", "strain")
#Then for Pa:
nonExpAllDataPa %>% filter(time=="12")->nonExpAllDataPaYield
nonExpAllDataPaYield %>% select(countsmL,rep,bug) %>% mutate(condition = "static")->nonExpAllDataPaYieldSimple
nonExpAllDataPaYieldSimple<-nonExpAllDataPaYieldSimple[c(1,2,4,3)]
colnames(nonExpAllDataPaYieldSimple) <- c("bugs","rep", "condition", "strain")

##Now get the growth yield for mono starting from exponential phase
#First for Sa:
yieldSaFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Pa growth yields 16 hrs/Sa growth yields 16 hrs.csv"
yieldSaData <-read.csv(yieldSaFile)
ggplot()+
  geom_point(data=yieldSaData, aes(x=condition, y=bugs), size=4)+
  scale_y_log10(limits= c(1e6,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(6,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste("Sa growth yield (16 hrs)")), x = "condtion")
#Then for Pa:
yieldPaFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Pa growth yields 16 hrs/Pa growth yields 16 hrs.csv"
yieldPaData <-read.csv(yieldPaFile)
yieldPaData<-yieldPaData %>% filter(!bio.rep==3)#third replicate was thrown out because plate counts were off
ggplot()+
  geom_jitter(data=yieldPaData, aes(x=condition, y=bugs, color = strain), size=4, height = 0, width = .1)+
  scale_y_log10(limits= c(1e6,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(6,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste("Pa growth yield (16 hrs)")), x = "condition")
```

Shown below are the combination of the 12th hr timepoint of non exponential growth curves and the yield after 16 hrs of exp growth.
```{r combined yield, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}
##now, combining both exponential and non-exponential, we get this:
#First for Sa:
allYieldsSa <- gdata::combine(yieldSaData, nonExpAllDataSaYieldSimple)
ggplot()+
  geom_jitter(data=allYieldsSa, aes(x=condition, y=bugs, color = source), size=4, height = 0, width = .1)+
  scale_y_log10(limits= c(1e6,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(6,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  scale_color_discrete(name = "Source",labels = c("Exp", "NonExp"))+
  labs(title =expression(paste("Sa Growth Yield Combined")), x = "condition")

allYieldsSa %>% group_by(condition) %>% summarise(meanYield = mean(bugs), SEYield = std.error(bugs)) %>% as.data.frame()->yieldsSaFinal

#Now for Pa:
yieldPaData<-yieldPaData %>% group_by(bio.rep, condition, strain) %>% summarise(bugsmL= mean(bugs))%>% as.data.frame() #data from Pa were not as simple as for Sa
yieldPaData<-yieldPaData[c(4,1,2,3)]
colnames(yieldPaData) <- c("bugs","rep", "condition", "strain")

allYieldsPa <- gdata::combine(yieldPaData, nonExpAllDataPaYieldSimple)
allYieldsPa<-allYieldsPa %>% mutate(strains = recode(allYieldsPa$strain, wt = "PA14wt", pqsL = "PA14-pqsL")) %>% as.data.frame()

ggplot()+
  geom_jitter(data=allYieldsPa, aes(x=condition, y=bugs, color = source, shape = strains), width = .1, height = 0,size=4)+
  scale_y_log10(limits= c(1e7,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(6,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  scale_color_discrete(name = "Source",labels = c("Exp", "NonExp"))+
  labs(title =expression(paste("Pa Growth Yield Combined")), x = "condition")

allYieldsPa %>% group_by(condition, strains) %>% summarise(meanYield = mean(bugs), SEYield = std.error(bugs)) %>% as.data.frame()->yieldsPaFinal
ggplot()+
  geom_jitter(data=allYieldsPa, aes(x=condition, y=bugs, color = source), width = .1, height = 0,size=4)+
  #geom_point(aes(x = condition, y-meanYield-SEYield, ymax=meanYield+SEYield), data =yieldsPaFinal, width=.1)+
  geom_errorbar(aes(x = condition, ymin=meanYield-SEYield, ymax=meanYield+SEYield), data =yieldsPaFinal, width=.1)+
  scale_y_log10(limits= c(1e8,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(6,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  labs(title =expression(paste("Pa Growth Yield Combined")), x = "condition")+
  scale_color_discrete(name = "Source",labels = c("Exp", "NonExp"))+
  facet_grid(.~strains)
```

This is the growth curve, followed by the plots with horizontal bars showing the average growth yield among all replicates.

```{r final plots, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}

ggplot()+
  geom_line(data=coCultureGCDataStats,aes(x=time, y=cellCount, color = bug, linetype = strain), size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD, color = bug), data =coCultureGCDataStats, width=.2, size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD,color = bug), data =coCultureGCDataStats, width=.2, size = 1)+
  scale_y_log10(limits= c(1e5,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(.~condition)+
  scale_x_continuous(breaks = c(0,4,8,12,16))+
  #scale_linetype_discrete(guide = FALSE)+
  labs(title =expression(paste(italic("P. aeruginosa"), " PA14 wt and ", italic("S. aureus"), " coculture in SCFM2")), x = "Time (hours)", shape = "Strain", color = "Species")

coCultureGCDataStatsMixed<-coCultureGCDataStats %>% filter(condition == "Stirred") %>% as.data.frame()
coCultureGCDataStatsMixedSa<-coCultureGCDataStatsMixed %>% filter(bug == "Sa") %>% as.data.frame()
coCultureGCDataStatsMixedPa<-coCultureGCDataStatsMixed %>% filter(bug == "Pa") %>% as.data.frame()

coCultureGCDataStatsStatic<-coCultureGCDataStats %>% filter(condition == "Static") %>% as.data.frame()
coCultureGCDataStatsStaticSa<-coCultureGCDataStatsStatic %>% filter(bug == "Sa") %>% as.data.frame()
coCultureGCDataStatsStaticPa<-coCultureGCDataStatsStatic %>% filter(bug == "Pa") %>% as.data.frame()

##Now we have everything we need for the final plot
#these are the yields:
yieldPaShMut<-yieldsPaFinal[1,3]
yieldPaShWt<-yieldsPaFinal[2,3]
yieldPaStMut<-yieldsPaFinal[3,3]
yieldPaStWt<-yieldsPaFinal[4,3]
yieldSaSt<-yieldsSaFinal[1,2]
yieldSaSh<-yieldsSaFinal[2,2]

#Plot for static data
ggplot()+
  geom_line(data=coCultureGCDataStatsStaticSa,aes(x=time, y=cellCount, linetype = strain), color = "red2", size = 1)+
  geom_line(data=coCultureGCDataStatsStaticPa,aes(x=time, y=cellCount, linetype = strain), color = "green3", size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "red2", data =coCultureGCDataStatsStaticSa, width=.2, size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "green3", data =coCultureGCDataStatsStaticPa, width=.2, size = 1)+
  scale_y_log10(limits= c(1e6,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  geom_hline(yintercept=yieldSaSt, linetype = "solid", color = "red2", size = 1)+
  geom_hline(yintercept=yieldPaStWt, linetype = "solid", color = "green3",size = 1)+
  geom_hline(yintercept=yieldPaStMut, linetype = "dashed", color = "green3",size = 1)+
  scale_x_continuous(breaks = c(0,4,8,12,16))+
  #scale_linetype_discrete(guide = FALSE)+
  labs(title =expression(paste(italic("P. aeruginosa"), " PA14 wt and ", italic("S. aureus"), " coculture static in SCFM2")), x = "Time (hours)", shape = "Strain", color = "Species")

ggplot()+
  geom_line(data=coCultureGCDataStatsMixedSa,aes(x=time, y=cellCount, linetype = strain), color = "red2", size = 1)+
  geom_line(data=coCultureGCDataStatsMixedPa,aes(x=time, y=cellCount, linetype = strain), color = "green3", size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "red2", data =coCultureGCDataStatsMixedSa, width=.2, size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "green3", data =coCultureGCDataStatsMixedPa, width=.2, size = 1)+
  scale_y_log10(limits= c(1e5,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  geom_hline(yintercept=yieldSaSh, linetype = "solid", color = "red2", size = 1)+
  geom_hline(yintercept=yieldPaShWt, linetype = "solid", color = "green3",size = 1)+
  geom_hline(yintercept=yieldPaShMut, linetype = "dashed", color = "green3",size = 1)+
  scale_x_continuous(breaks = c(0,4,8,12,16))+
  #scale_linetype_discrete(guide = FALSE)+
  labs(title =expression(paste(italic("P. aeruginosa"), " PA14 wt and ", italic("S. aureus"), " coculture stirred in SCFM2")), x = "Time (hours)", shape = "Strain", color = "Species")

```
And shown below are the same graphs but only using the data from exp phase growth yields
```{r final plots exp only, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}

yieldSaData %>% group_by(condition) %>% summarise(meanYield = mean(bugs), SEYield = std.error(bugs)) %>% as.data.frame()->yieldsSaFinalExpOnly

yieldPaData %>% group_by(condition, strain) %>% summarise(meanYield = mean(bugs), SEYield = std.error(bugs)) %>% as.data.frame()->yieldsPaFinalExpOnly

yieldPaShMutExp<-yieldsPaFinalExpOnly[1,3]
yieldPaShWtExp<-yieldsPaFinalExpOnly[2,3]
yieldPaStMutExp<-yieldsPaFinalExpOnly[3,3]
yieldPaStWtExp<-yieldsPaFinalExpOnly[4,3]
yieldSaStExp<-yieldsSaFinalExpOnly[1,2]
yieldSaShExp<-yieldsSaFinalExpOnly[2,2]

ggplot()+
  geom_line(data=coCultureGCDataStatsStaticSa,aes(x=time, y=cellCount, linetype = strain), color = "red2", size = 1)+
  geom_line(data=coCultureGCDataStatsStaticPa,aes(x=time, y=cellCount, linetype = strain), color = "green3", size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "red2", data =coCultureGCDataStatsStaticSa, width=.2, size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "green3", data =coCultureGCDataStatsStaticPa, width=.2, size = 1)+
  scale_y_log10(limits= c(1e6,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  geom_hline(yintercept=yieldSaStExp, linetype = "solid", color = "red2", size = 1)+
  geom_hline(yintercept=yieldPaStWtExp, linetype = "solid", color = "green3",size = 1)+
  geom_hline(yintercept=yieldPaStMutExp, linetype = "dashed", color = "green3",size = 1)+
  scale_x_continuous(breaks = c(0,4,8,12,16))+
  #scale_linetype_discrete(guide = FALSE)+
  labs(title =expression(paste(italic("P. aeruginosa"), " PA14 wt and ", italic("S. aureus"), " coculture static in SCFM2 (only exp yield)")), x = "Time (hours)", shape = "Strain", color = "Species")

ggplot()+
  geom_line(data=coCultureGCDataStatsMixedSa,aes(x=time, y=cellCount, linetype = strain), color = "red2", size = 1)+
  geom_line(data=coCultureGCDataStatsMixedPa,aes(x=time, y=cellCount, linetype = strain), color = "green3", size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "red2", data =coCultureGCDataStatsMixedSa, width=.2, size = 1)+
  geom_errorbar(aes(x = time, ymin=cellCount-cellSD, ymax=cellCount+cellSD), color = "green3", data =coCultureGCDataStatsMixedPa, width=.2, size = 1)+
  scale_y_log10(limits= c(1e5,1e10), name=expression(paste('CFUs per mL' )), breaks = c(10^(seq(5,10,.5))), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  geom_hline(yintercept=yieldSaShExp, linetype = "solid", color = "red2", size = 1)+
  geom_hline(yintercept=yieldPaShWtExp, linetype = "solid", color = "green3",size = 1)+
  geom_hline(yintercept=yieldPaShMutExp, linetype = "dashed", color = "green3",size = 1)+
  scale_x_continuous(breaks = c(0,4,8,12,16))+
  #scale_linetype_discrete(guide = FALSE)+
  labs(title =expression(paste(italic("P. aeruginosa"), " PA14 wt and ", italic("S. aureus"), " coculture stirred in SCFM2 (only exp yield)")), x = "Time (hours)", shape = "Strain", color = "Species")
```

```{r killing assays, echo=FALSE, warning=FALSE,fig.width=9, fig.height=7}
diskFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Staph killing assays/disk difussion assays.csv"
diskData <-read.csv(diskFile)
diskData<-diskData %>% mutate (zoneInhibition = avg-6)
diskData%>% group_by(condition, strain) %>% summarise(meanZone = mean(zoneInhibition), SEZone = std.error(zoneInhibition)) %>% as.data.frame() -> diskDataStats

ggplot()+#both PA and Sa
  geom_point(data=diskData, aes(x=condition, y=zoneInhibition, shape = factor(tech.rep), color = factor(bio.rep)), size=4)+
  facet_grid(.~strain)+
  labs(title =expression(paste("Zone of inhibition")), x = "condtion",shape = "tech rep", color = "bio rep") 

ggplot()+#both PA and Sa
  geom_point(data=diskDataStats, aes(x=condition, y=meanZone, color =  strain), size=4)+
  geom_errorbar(aes(x = condition, ymin=meanZone-SEZone, ymax=meanZone+SEZone, color = strain), data =diskDataStats, width=.2, size = 1)+
  labs(title =expression(paste("Zone of inhibition")), x = "condtion",shape = "tech rep", color = "bio rep") 

```